Manuscript

### 2025-02-06 Manuscript_MD_ScoreReLiability.qmd
update MS

@@

Large-scale collaborative replication efforts have sparked discussions
surrounding the replicability of psychological phenomena. However,
attempts to estimate a phenomenon’s replicability or to predict whether
a future replication will replicate successfully, heterogeneity is a
crucial parameter that needs to be assessed. In the meta-analytic
context, heterogeneity describes the variation in effect sizes, free of
sampling error, if these are not stable or constant across replications.
Therefore, the presence of heterogeneity implies that some replications
of a single phenomenon may be successful, while others are not. If
heterogeneity grows, meaning the phenomenon’s effect size varies more
strongly for unexplained reasons, the probability of observing an effect
size around zero or even in the negative space grows larger as well. If
we know the mean size and heterogeneity of a phenomenon’s effect size,
we would theoretically be able to establish a baseline of expected
replication rate (reference).

@@

In both the initial reports of large scale replication attempts, as well
as the re-analyses by Olsson-Collentine et al. (2020), standardized
effect sizes (from here-on abbreviated as ES) such as Cohen’s d or
Hedge’s g were used. For the remainder of the article, Cohen’s d,
defined in Equation 1, will be used as an exemplary estimate of ES, as
it is used across a wide range of contexts and well understood by a
broad audience.

$$d = \frac{MD}{\sigma_X}$${#eq-d}

Here, MD refers to the difference between two groups of interest, while
$\sigma_X$ is the total pooled standard deviation. Since $d$ constitutes
an observed effect size it is affected by sampling error. In terms of
heterogeneity, we are not interested in the variance of $d$, but the
variance of the true, underlying ES $\delta$. The standard error of
Cohen’s $d$, which is used to weight individual estimates in a
random-effects meta-analysis (reference) is computed as defined in
Equation 2.

$$SE_d=\sqrt{\frac{n_1+n_2}{n_1n_2}+\frac{d^2}{2\left(n_1+n_2\right)}}$${#eq-SEd}

As is widely known, score reliability affects ES. In the context of
classical test theory (CTT) score reliability ρXX’ is defined as the
ratio of true to total score variance, as defined in Equation 3.

$$\rho_{XX’} = \frac{\sigma^2_T}{\sigma^2_X}=\frac{\sigma^2_T}{\sigma^2_T + \sigma^2_E}$${#eq-rel}

@@

If information concerning score reliability is
available, it is possible to correct the individual ES for its
unreliability. This process is also known as attenuation correction and,
while not without its criticisms (reference), is a widespread practice
in Psychology (reference). Equation 4 describes a simple attenuation
correction procedure.

$$d_c = \frac{d}{\sqrt{\hat{\rho}_{xx’}}}$${#eq-dc}

Here, $d_c$ describes the attenuation-corrected estimate of $d$, which
is corrected by dividing it by the estimate of score reliability
$\hat{\rho}_{xx’}$. As described in W&D or Lord & Novick (references),
this is essentially the same, as if the corrected estimate of $d_c$ was
constructed using an estimate of true score standard deviation, as in
Equation 5.

$$d_c = \frac{MD}{\hat{\sigma}_{T}}=\frac{MD}{\sqrt{\hat{\rho}_{xx’} \hat{\sigma}_X^2}}$${#eq-dct}

$\hat{\sigma}_T$ can be estimated by simply rearranging the terms from
equation 2. Thereby, an estimate of ES is generated, under the premise
that score reliability would have been perfect ($\rho_{xx'} = 1$). As
the corrected estimate of Cohen’s $d$ is larger and requires an
additional unknown parameter, an estimate of score reliability, the
larger uncertainty in this parameter should be reflected in a larger
standard error. The standard error for $d_c$ is defined in equation 6.

$$SE_{d_c} = \sqrt{SE^2_d \frac{d_c}{d}}$${#eq-SEdc}

@@

Re-analyses of the collaborative multi-site replication efforts have
uncovered differences in score reliability across replications of the
same phenomena (ref McShaw etc.). The claims made in W&D and H&S imply
that the ES heterogeneity uncovered in these projects therefore could
potentially be explained by these differences in score reliability. The
same argument has been proposed by both original authors of phenomena
that did not replicate or carry substantial heterogeneity (references)
as well as the authors of the multi-site replication attempts (refs).
Over the following pages, we assess for the first time, whether the
differences in score reliability do indeed explain, and thereby reduce,
the ES heterogeneity observed across the different multi-site
replication efforts. Even though only a limited number of phenomena
allow for such an analysis, we find no evidence for a reduction in ES
heterogeneity after taking score reliability into account. Subsequently,
we discuss why the expectation formulated in H&S does not allow for such
claims. Alternatively, we explore to what extent defining ES as a random
ratio variable helps understand how ES heterogeneity is affected by
differences in score reliability. We close with a discussion on what
future meta-analysts can expect to observe, when dealing with issues of
(un)reliability in assessments of meta-analyses of ES.

## Re-analysis of archival data Data

@@

@@

While replication data on a decent number of psychological phenomena is
available in the DRIPHT-repository, to demonstrate how incorporating
score reliability affects heterogeneity in ES, selected phenomena need
to fulfil three conditions: (1) As we focus on the use of ES $d$ d, the
phenomenon needs to be assessed using a two-group design. This is the
case for all projects in the DRIPHT repository; (2) Score reliability
estimates can be derived, this implies that the phenomenon needs to be
measured using either several indicators forming a single scale or by
repeatedly measuring across several timepoints; (3) Score reliability
can only attenuate effect sizes that are not zero in the first place.
Therefore, we focus on phenomena where meta-analytic mean ES can be
statistically distinguished from zero.

@@

For each replication of a phenomenon, Cohen’s $d$ is computed as an
estimate of raw ES, according to Equation 1, including its standard
error (Equation 2). Subsequently, a random-effects meta-analysis is
performed using metafor version X.XX. To estimate meta-analytic mean ES
and heterogeneity, the REML-estimatore is used (ref). To identify which
phenomena pass criterion (3), using a Wald-type significance test we
assess for which phenomenon the meta-analytic mean ES is significantly
different from zero. For this hypothesis test, as all other hypothesis
tests in this manuscript, a significance level of .05 is used.

Additionally, estimates of score reliability are derived. While not
without its criticism (references), Cronbach’s Alpha is used to estimate
score reliability. It is often noted that tau-equivalence is an
unrealistic assumption to hold against real-world data, implying that
other estimates of score reliability, such as McDonald’s Omega
(reference), Guttman’s Lambda 2 or 4 (reference), or the Greatest Lower
Bound (reference) might be better suited. However, all scales employed
in this project are evaluated by computing the simple mean in responses
across items for each individual. Computing the mean of several items
implicitly assumes tau-equivalence, as each item contributes equally to
the individual’s test score. Therefore, Cronbach’s Alpha is the better
choice to estimate score reliability, as it avoids this mismatch in
assumptions between how the score is computed and how score reliability
is estimated. Using these estimates of Cronbach’s Alpha, the individual
estimates of Cohen’s d are corrected, according to Equation 4, including
the corrected standard errors (Equation 5). Thereby, estimates of ES are
computed, which are corrected for imperfect reliability. Again, a
random-effects meta-analysis is run using metafor, generating an
estimate of heterogeneity in corrected ES. Different indicators of
heterogeneity are readily available, such as $I^2$, $H^2$ or the
coefficient of variation. However, the descriptions found in H&S and W&D
discuss the absolute amount of heterogeneity in terms of variance
($\tau^2$) or standard deviation ($\tau$). These parameters discuss the
variability of the standardized ES in the population in terms of
variance or standard deviation. Therefore, we also make use of absolute
heterogeneity here.

@@

$$\mu_{\rho{xx'}} \approx 1 - exp(\mu_{T\left[\rho_{XX^\prime}\right]})-\frac{1}{2} exp(\mu_{T\left[\rho_{xx'}\right]})\tau_{T\left[\rho_{XX^\prime}\right]}^2$${#eq-rel_back_mu}

$$\tau_{\rho_{xx'}}^2 \approx exp(\mu_{T\left[\rho_{xx'}\right]})2 \tau_{T\left[\rho_{xx'}\right]}^2+exp(\mu_{T\left[\rho_{xx'}\right]})^2\tau_{T\left[\rho_{xx'}\right]}^4$${#eq-rel_back_tau}

@@

Equations 1, 2, 4, and 5 were used to generate estimates of standardized
ES, both uncorrected and corrected, with their corresponding standard
errors. Generally, this leads to larger (absolute) ES and larger
standard errors. As an example from the 12 selected phenomena, the
results from this procedure on Nosek’s Explicit Art sex differences
phenomenon are displayed in the forest plot in Figure 1. Here, grey dots
represent the uncorrected ES for each sample, while black dots represent
the corrected ES in each sample. The bars surrounding these dots show
the respective 95%-Cis.

In Figure 1 it becomes apparent that the reliability attenuation
procedure leads to an increase in the individual absolute effect size,
as the black dots are moved further away from zero. Simultaneously, the
standard errors grew larger after the attenuation correction, which
leads to larger 95%-Cis. This is most easily observed for the
mturk-sample, which already had a rather large confidence interval to
begin with. Additionally, in the diamond at the bottom of the figure, we
can see that the meta-analytic estimate mean ES is also larger after the
attenuation correction took place. In Table 2, the estimates and tests
on the meta-analytic average ES, both corrected and uncorrected, can be
found.

*Table 2: Meta-analytic average  ES* \| ------- \| ------- \|
------- \| ------- \| ------- \| \| MASC \|
{\hat{\mathbit{\mu}}}*{*\mathbit{\delta}\mathbit{o}} \| p \|
{\hat{\mathbit{\mu}}}\_\mathbit{\delta} \| p \| \|
Albarracin_Priming_SAT \| .127 (.053) \| .016 \| .158 (.066) \| .016 \|
\| Alter_Analytic_Processing \| -.155 (.040) \| \<.001 \| -.174 (.044)
\| \<.001 \| \| Carter_Flag_Priming \| .018 (.029) \| .538 \| .026
(.036) \| .465 \| \| Caruso_Currency_Priming \| -.019 (.026) \| .457 \|
-.022 (.029) \| .459 \| \| Finkel_Exit_Forgiveness \| -.050 (.049) \|
.314 \| -.056 (.056) \| .316 \| \| Finkel_Neglect_Forgiveness \| -.051
(.055) \| .355 \| -.059 (.063) \| .347 \| \| Giessner_Vertical_Position
\| .026 (.023) \| .248 \| .027 (.025) \| .267 \| \|
Hart_Criminal_Intentionality \| .170 (.078) \| .030 \| .184 (.085) \|
.030 \| \| Hart_Detailed_Processing \| .082 (.058) \| .157 \| .096
(.067) \| .147 \| \| Hart_Intention_Attribution \| -.011 (.058) \| .856
\| -.021 (.080) \| .795 \| \| Husnu_Imagined_Contact \| .116 (.032) \|
\<.001 \| .129 (.035) \| \<.001 \| \| Nosek_Explicit_Art \| .360 (.047)
\| \<.001 \| .383 (.050) \| \<.001 \| \| Nosek_Explicit_Math \| .401
(.029) \| \<.001 \| .415 (.030) \| \<.001 \| \| PSACR001_anxiety_int \|
.251 (.022) \| \<.001 \| .263 (.023) \| \<.001 \| \| PSACR001_behav_int
\| .035 (.020) \| .079 \| .040 (.023) \| .088 \| \| PSACR002_neg_photo
\| -.663 (.040) \| \<.001 \| -.747 (.047) \| \<.001 \| \|
Shnabel_Willingness_Reconcile_Rev \| -.533 (.101) \| \<.001 \| -.582
(.109) \| \<.001 \| \| Shnabel_Willingness_Reconcile_RPP \| -.585 (.062)
\| \<.001 \| -.655 (.067) \| \<.001 \| \| Srull_Behaviour_Hostility \|
-.051 (.029) \| .079 \| -.062 (.034) \| .071 \| \|
Srull_Ronald_Hostility \| .063 (.029) \| .028 \| .074 (.033) \| .026 \|
\| Tversky_Directionality_Similarity1 \| -.111 (.034) \| .001 \| -.121
(.037) \| \<.001 \| \| Zhong_Desirability_Cleaning \| -.031 (.024) \|
.194 \| -.042 (.036) \| .240 \|

Table 2 demonstrates that what we observed in Figure 1 holds across all
22 phenomena. All meta-analytic effect sizes are larger after applying
an attenuation correction procedure. Similarly, the uncertainty
quantified in the estimates’ standard error is larger. Additionally,
Table 2 highlights which phenomena pass criterion (3) – the effect size
must be statistically significantly distinguishable from zero. 12
phenomena pass this criterion, highlighted in boldface. It may be
remarked that the attenuation correction procedure, in this data,
appears to have no influence on whether a phenomenon passes the
criterion. The differences in p-value are rather small and lead to no
difference in conclusion, regarding a significance level of .05.

*Table 3: Tests for ES Heterogeneity* \| \| Corrected ES \| \|
Corrected ES \| NA \| NA \| NA \| NA \| \| ------- \| ------- \| -------
\| ------- \| ------- \| ------- \| ------- \| ------- \| \| MASC \|
$\tau$ \| QE \| p \| \| $\tau$ \| QE \| p \| \| Albarracin_Priming_SAT
\| 0 \| 5.987 (8) \| .649 \| \| 0 \| 6.055 (8) \| .641 \| \|
Alter_Analytic_Processing \| .025 \| 15.706 (20) \| .735 \| \| .018 \|
15.701 (20) \| .735 \| \| Hart_Criminal_Intentionality \| .171 \| 18.96
(11) \| .062 \| \| .185 \| 18.81 (11) \| .065 \| \|
Husnu_Imagined_Contact \| .087 \| 47.581 (35) \| .076 \| \| .094 \|
47.351 (35) \| .079 \| \| Nosek_Explicit_Art \| .178 \| 62.771 (35) \|
.003 \| \| .19 \| 63.074 (35) \| .003 \| \| Nosek_Explicit_Math \| 0 \|
39.169 (35) \| .288 \| \| 0 \| 39.107 (35) \| .29 \| \|
PSACR001_anxiety_int \| .085 \| 74.655 (48) \| .008 \| \| .088 \| 74.251
(48) \| .009 \| \| PSACR002_neg_photo \| .22 \| 219.709 (36) \| \<.001
\| \| .26 \| 237.115 (36) \| \<.001 \| \| Shnabel\_ Reconcile_Rev \|
.233 \| 23.367 (7) \| .001 \| \| .253 \| 23.182 (7) \| .002 \| \|
Shnabel\_ Reconcile_RPP \| .073 \| 8.472 (7) \| .293 \| \| .071 \| 8.244
(7) \| .312 \| \| Srull_Ronald_Hostility \| .046 \| 24.756 (21) \| .258
\| \| .053 \| 24.593 (21) \| .265 \| \| Tversky\_ Similarity1 \| .001 \|
59.434 (60) \| .496 \| \| .001 \| 59.102 (60) \| .509 \|

@@

However, what Figure 1 and Table 2 can not inform us about is in how far
the heterogeneity has changed after correcting for (un)reliability. In
Table 3, the estimates and tests of heterogeneity concerning uncorrected
and corrected ES on the 12 remaining phenomena, which passed all
criteria (1) – (3), can be found. In the table, the estimate of
heterogeneity $\tau$, with the accompanying QE-test statistic, degrees
of freedom and its p-value are reported, separately for uncorrected and
corrected ES.

Most importantly, Table 3 demonstrates that of 12 phenomena assessed,
for seven phenomena we find patterns incompatible with claims made in
W&D and H&S. For those seven phenomena, the reliability attenuation
correction led to an increase in absolute ES heterogeneity $\tau$. For
the remaining five phenomena, estimates of ES heterogeneity were zero
before and after the correction procedure, leading to no change at all.
This leaves three phenomena where the attenuation correction did indeed
lead to a reduction of heterogeneity. However, it is crucial to note
that the hypothesis tests identified statistically significant
heterogeneity in only four out of 12 phenomena and for none of those a
reduction in heterogeneity took place. In all four cases of
statistically significant heterogeneity, the attenuation correction
procedure led to an increase in heterogeneity, contrary to the W&D and
H&S predictions.

To make sure that actual differences in score reliability exist,
heterogeneity in score reliability is assessed, making use of
Reliability Generalization Meta-Analysis (reference). Table 4 summarises
the results across all 12 phenomena. Overall, scales for almost all
phenomena appear have produced scores with an average score reliability
larger than .7. Only two phenomena (Albarracin_Priming_SAT and
Alter_Analytic_Processing) come with lower score reliability. However,
for those phenomena we could not identify statistically significant
heterogeneity in effect sizes in Table 2 in the first place. From the
phenomena where the ES heterogeneity grew larger as a result from the
attenuation correction procedure, for six out of seven phenomena, we
observe statistically significant heterogeneity in score reliability.
Only the Shnabel_Willingness_Reconcile_Rev project produced scores where
no heterogeneity in score reliability was identified. However, for this
project, it is crucial to note that only a small number of replications
(8), was available for analysis. Since the power of Cochran’s Q-test for
heterogeneity largely depends on the number of replications, this result
is inconclusive.

@@

*Table 4: Reliability Generalization Meta-Analysis MASC* \| MASC \|
\mathbit{\mu}*{*\mathbit{\rho}{\mathbit{xx}\prime}} \|
\mathbit{\tau}*{*\mathbit{\rho}{\mathbit{xx}\prime}} \| QE \| p \| \|
------- \| ------- \| ------: \| ------- \| ------- \| \|
Albarracin_Priming_SAT \| .61 \[.557 : .657\] \| .057 \| 26.019 (9) \|
.001 \| \| Alter_Analytic_Processing \| .64 \[.604 : .673\] \| .056 \|
38.384 (20) \| .008 \| \| Hart_Criminal_Intentionality \| .849 \[.811 :
.879\] \| .054 \| 66.17 (11) \| \<.001 \| \| Husnu_Imagined_Contact \|
.814 \[.802 : .825\] \| .023 \| 66.389 (35) \| .001 \| \|
Nosek_Explicit_Art \| .881 \[.869 : .891\] \| .029 \| 142.551 (35) \|
\<.001 \| \| Nosek_Explicit_Math \| .936 \[.932 :.939\] \| .007 \| 59.43
(35) \| .006 \| \| PSACR001_anxiety_int \| .903 \[.89 : .914\] \| .041
\| 494.822 (48) \| \<.001 \| \| PSACR002_neg_photo \| .791 \[.774 :
.806\] \| .047 \| 350.852 (36) \| \<.001 \| \|
Shnabel_Willingness_Reconcile_Rev \| .832 \[.818 : .845\] \| 0 \| 5.053
(7) \| .653 \| \| Shnabel_Willingness_Reconcile_RPP \| .768 \[.737:
.796\] \| .032 \| 17.535 (7) \| .014 \| \| Srull_Ronald_Hostility \|
.739 \[.716 : .76\] \| .047 \| 91.722 (21) \| \<.001 \| \|
Tversky_Directionality_Similarity1 \| .854 \[.841 : .866\] \| .039 \|
164.933 (60) \| \<.001 \|

@@

$$\tau^2_{\delta_0} = [\mu_{\rho_{xx‘}}]^2 \tau^2_{\delta} + [\mu_{\delta}]^2 \tau^2_{\rho_{xx‘}}$${#eq-HS}

We adjusted Equation 9 using the notation used throughout this text.
Here, $\delta_0$ refers to the ES, undistorted by sampling error but not
corrected for measuring error, $\delta$ refers to the ES, undistorted by
sampling error and measurement error and ρxx‘ refers to score
reliability.

Equation 9 implies that heterogeneity in $\delta_0$ is essentially a
function where heterogeneity and mean value of $\delta$ are weighted by
mean score reliability ρxx‘ and its heterogeneity. While it may not be
self-evident in the equation itself, it is easy to construct cases where
heterogeneity in $\delta_0$ is inflated or deflated by heterogeneity in
ρxx‘, contradicting Hunter & Schmidt’s interpretation of this equation.
In Figure 2, the results of employing equation 9 to compute the
heterogeneity (variance) in corrected ES is highlighted. All else held
equal, only the individual level of mean reliability \$
\mu*{*\rho{xx‘}}\$ is varied from .5 to .8. Mean corrected ES
$\mu_{\delta}$ is held constant at 1, its variance $\tau^2_{\delta}$
held constant at .04 and the variance of score reliability
$\tau^2_{\rho_{xx‘}}$ held constant at .02.

*Figure 2*

In Figure 2, the red line corresponds to the variance of corrected ES.

>>>>

```{r, include = FALSE}
# library loading and installing as necessary

# relevant libraries required for this script
packages <- c("ggplot2", "here", "metafor", "gridExtra", "knitr", "magrittr", "dplyr")

# check, whether library already installed or not - install and load as needed:
apply(as.matrix(packages), MARGIN = 1, FUN = function(x) {

  pkg_avail <- nzchar(system.file(package = x))   # check if library is installed on system

  if(pkg_avail){
    require(x, character.only = TRUE)             # load the library, if already installed

  }else{
    install.packages(x, repos = "https://ftp.gwdg.de/pub/misc/cran/")                           # install the library, if missing
    require(x, character.only = TRUE)             # load after installation
  }
})

source(here("ReLiability_Function-library.R"))

ES_rma_df <- read.csv(here("Data/Processed/Aggregates_ES_analysis.csv"))

agg_L <- readRDS(here("Data/Processed/Aggregates_simple.csv"))

MASC_names <- unique(ES_rma_df$MASC)

```

@@

Large-scale collaborative replication efforts have sparked discussions
surrounding the replicability of psychological phenomena. However, for
attempts to estimate a phenomenon’s replicability or to predict whether
a future replication will replicate successfully, heterogeneity is a
crucial parameter that needs to be assessed. In the meta-analytic
context, heterogeneity describes the variation in effect sizes, free of
sampling error, if these are not stable or constant across replications.
Therefore, the presence of heterogeneity implies that some replications
of a single phenomenon may be successful, while others are not. If
heterogeneity grows, meaning the phenomenon’s effect size varies more
strongly for unexplained reasons, the probability of observing an effect
size around zero or even in the negative space grows larger as well. If
we know the mean size and heterogeneity of a phenomenon’s effect size,
we would theoretically be able to establish a baseline of expected
replication rate (reference).

@@

In both the initial reports of large scale replication attempts, as well
as the re-analyses by Olsson-Collentine et al. (2020), standardized
effect sizes (from here-on abbreviated as ES) such as Cohen’s d or
Hedge’s g were used. For the remainder of the article, Cohen’s d,
defined in @eq-d, will be used as an exemplary estimate of ES, as it is
used across a wide range of contexts and well understood by a broad
audience.

$$d = \frac{MD}{\sigma_X}$$ {#eq-d}

Here, MD refers to the difference between two groups of interest, while
$\sigma_X$ is the total pooled standard deviation. Since $d$ constitutes
an observed effect size it is affected by sampling error. In terms of
heterogeneity, we are not interested in the variance of $d$, but the
variance of the true, underlying ES $\delta$. The standard error of
Cohen’s $d$, which is used to weight individual estimates in a
random-effects meta-analysis (reference) is computed as defined in
@eq-SEd.

$$SE_d=\sqrt{\frac{n_1+n_2}{n_1n_2}+\frac{d^2}{2\left(n_1+n_2\right)}}$$ {#eq-SEd}

As is widely known, score reliability affects ES. In the context of
classical test theory (CTT) score reliability $\rho_{XX’}$ is defined as
the ratio of true to total score variance, as defined in @eq-rel.

$$\rho_{XX’} = \frac{\sigma^2_T}{\sigma^2_X}=\frac{\sigma^2_T}{\sigma^2_T + \sigma^2_E}$$ {#eq-rel}

@@

If information concerning score reliability is available, it is possible
to correct the individual ES for its unreliability. This process is also
known as attenuation correction and, while not without its criticisms
(reference), is a widespread practice in Psychology (reference). @eq-dc
describes a simple attenuation correction procedure.

$$d_c = \frac{d}{\sqrt{\hat{\rho}_{xx’}}}$$ {#eq-dc}

Here, $d_c$ describes the attenuation-corrected estimate of $d$, which
is corrected by dividing it by the estimate of score reliability
$\hat{\rho}_{xx’}$. As described in W&D or Lord & Novick (references),
this is essentially the same, as if the corrected estimate of $d_c$ was
constructed using an estimate of true score standard deviation, as in
@eq-dct.

$$d_c = \frac{MD}{\hat{\sigma}_{T}}=\frac{MD}{\sqrt{\hat{\rho}_{xx’} \hat{\sigma}_X^2}}$$ {#eq-dct}

$\hat{\sigma}_T$ can be estimated by simply rearranging the terms from
@eq-rel. Thereby, an estimate of ES is generated, under the premise that
score reliability would have been perfect ($\rho_{xx'} = 1$). As the
corrected estimate of Cohen’s $d$ is larger and requires an additional
unknown parameter, an estimate of score reliability, the larger
uncertainty in this parameter should be reflected in a larger standard
error. The standard error for $d_c$ is defined in @eq-SEdc.

$$SE_{d_c} = \sqrt{SE^2_d \frac{d_c}{d}}$$ {#eq-SEdc}

@@

Re-analyses of the collaborative multi-site replication efforts have
uncovered differences in score reliability across replications of the
same phenomena (ref McShaw etc.). The claims made in W&D and H&S imply
that the ES heterogeneity uncovered in these projects therefore could
potentially be explained by these differences in score reliability. The
same argument has been proposed by both original authors of phenomena
that did not replicate or carry substantial heterogeneity (references)
as well as the authors of the multi-site replication attempts (refs).
Over the following pages, we assess for the first time, whether the
differences in score reliability do indeed explain, and thereby reduce,
the ES heterogeneity observed across the different multi-site
replication efforts. Even though only a limited number of phenomena
allow for such an analysis, we find no evidence for a reduction in ES
heterogeneity after taking score reliability into account. Subsequently,
we discuss why the expectation formulated in H&S does not allow for such
claims. Alternatively, we explore to what extent defining ES as a random
ratio variable helps understand how ES heterogeneity is affected by
differences in score reliability. We close with a discussion on what
future meta-analysts can expect, when dealing with issues of
(un)reliability in assessments of meta-analyses of ES.

## Re-analysis of archival data

@@

### Data

@@

While replication data on a decent number of psychological phenomena is
available in the DRIPHT-repository, to demonstrate how incorporating
score reliability affects heterogeneity in ES, selected phenomena need
to fulfil three conditions: (1) As we focus on the use of ES $d$, the
phenomenon needs to be assessed using a two-group design. This is the
case for all projects in the DRIPHT repository; (2) Score reliability
estimates can be derived, this implies that the phenomenon needs to be
measured using either several indicators forming a single scale or by
repeatedly measuring across several timepoints; (3) Score reliability
can only attenuate effect sizes that are not zero in the first place.
Therefore, we focus on phenomena where meta-analytic mean ES can be
statistically distinguished from zero.

@@

For each replication of a phenomenon, Cohen’s $d$ is computed as an
estimate of raw ES, according to @eq-d, including its standard error
(@eq-SEd). Subsequently, a random-effects meta-analysis is performed
using metafor version X.XX. To estimate meta-analytic mean ES and
heterogeneity, the REML-estimatore is used (ref). To identify which
phenomena pass criterion (3), using a Wald-type significance test we
assess for which phenomenon the meta-analytic mean ES is significantly
different from zero. For this hypothesis test, as all other hypothesis
tests in this manuscript, a significance level of .05 is used.

Additionally, estimates of score reliability are derived. While not
without its criticism (references), Cronbach’s Alpha is used to estimate
score reliability. It is often noted that tau-equivalence is an
unrealistic assumption to hold against real-world data, implying that
other estimates of score reliability, such as McDonald’s Omega
(reference), Guttman’s Lambda 2 or 4 (reference), or the Greatest Lower
Bound (reference) might be better suited. However, all scales employed
in this project are evaluated by computing the simple mean in responses
across items for each individual. Computing the mean of several items
implicitly assumes tau-equivalence, as each item contributes equally to
the individual’s test score. Therefore, Cronbach’s Alpha is the better
choice to estimate score reliability, as it avoids this mismatch in
assumptions between how the score is computed and how score reliability
is estimated. Using these estimates of Cronbach’s Alpha, the individual
estimates of Cohen’s d are corrected, according to @eq-dc, including the
corrected standard errors (@eq-SEdc). Thereby, estimates of ES are
computed, which are corrected for imperfect reliability. Again, a
random-effects meta-analysis is run using metafor, generating an
estimate of heterogeneity in corrected ES. Different indicators of
heterogeneity are readily available, such as $I^2$, $H^2$ or the
coefficient of variation. However, the descriptions found in H&S and W&D
discuss the absolute amount of heterogeneity in terms of variance
($\tau^2$) or standard deviation ($\tau$). These parameters discuss the
variability of the standardized ES in the population in terms of
variance or standard deviation. Therefore, we also make use of absolute
heterogeneity here.

@@

$$\mu_{\rho{xx'}} \approx 1 - exp(\mu_{T\left[\rho_{XX^\prime}\right]})-\frac{1}{2} exp(\mu_{T\left[\rho_{xx'}\right]})\tau_{T\left[\rho_{XX^\prime}\right]}^2$$ {#eq-rel_back_mu}

$$\tau_{\rho_{xx'}}^2 \approx exp(\mu_{T\left[\rho_{xx'}\right]})2 \tau_{T\left[\rho_{xx'}\right]}^2+exp(\mu_{T\left[\rho_{xx'}\right]})^2\tau_{T\left[\rho_{xx'}\right]}^4$$ {#eq-rel_back_tau}

@@

[Equations @eq-d], [-@eq-SEd], [-@eq-dc], and [-@eq-SEdc] were used to
generate estimates of standardized ES, both uncorrected and corrected,
with their corresponding standard errors. Generally, this leads to
larger (absolute) ES and larger standard errors. As an example from the
12 selected phenomena, the results from this procedure on Nosek’s
Explicit Art sex differences phenomenon are displayed in the forest plot
in Figure 1. Here, grey dots represent the uncorrected ES for each
sample, while black dots represent the corrected ES in each sample. The
bars surrounding these dots show the respective 95%-Cis.

In @fig-forest it becomes apparent that the reliability attenuation
procedure leads to an increase in the individual absolute effect size,
as the black dots are moved further away from zero. Simultaneously, the
standard errors grew larger after the attenuation correction, which
leads to larger 95%-Cis. This is most easily observed for the
mturk-sample, which already had a rather large confidence interval to
begin with. Additionally, in the diamond at the bottom of the figure, we
can see that the meta-analytic estimate mean ES is also larger after the
attenuation correction took place. In @tbl-meanES, the estimates and
tests on the meta-analytic average ES, both corrected and uncorrected,
can be found.

```{r, echo = FALSE}
#| label: fig-forest
#| fig-cap: Forest Plot Nosek_Explicit_Art
forest_plot_rel(ES_rma_df[which(ES_rma_df$MASC == "Nosek_Explicit_Art"),], agg_L[[which(MASC_names == "Nosek_Explicit_Art")]])
```

```{r, echo = FALSE}
#| label: tbl-meanES
#| tbl-cap: Meta-analytic average  ES
knitr::kable(x = read.csv(here("Tables/Mean_ES.csv")))
```

@tbl-meanES demonstrates that what we observed in Figure 1 holds across
all 22 phenomena. All meta-analytic effect sizes are larger after
applying an attenuation correction procedure. Similarly, the uncertainty
quantified in the estimates’ standard error is larger. Additionally,
@tbl-meanES highlights which phenomena pass criterion (3) – the effect
size must be statistically significantly distinguishable from zero. 12
phenomena pass this criterion, highlighted in boldface. It may be
remarked that the attenuation correction procedure, in this data,
appears to have no influence on whether a phenomenon passes the
criterion. The differences in p-value are rather small and lead to no
difference in conclusion, regarding a significance level of .05.

```{r, echo = FALSE}
#| label: tbl-tauES
#| tbl-cap: Tests for ES Heterogeneity
knitr::kable(x = read.csv(here("Tables/Heterogeneity_ES.csv")))
```

@@

However, what @fig-forest and @tbl-meanES can not inform us about is in
how far the heterogeneity has changed after correcting for
(un)reliability. In @tbl-tauES, the estimates and tests of heterogeneity
concerning uncorrected and corrected ES on the 12 remaining phenomena,
which passed all criteria (1) – (3), can be found. In the table, the
estimate of heterogeneity $\tau$, with the accompanying QE-test
statistic, degrees of freedom and its p-value are reported, separately
for uncorrected and corrected ES.

Most importantly, @tbl-tauES demonstrates that of 12 phenomena assessed,
for seven phenomena we find patterns incompatible with claims made in
W&D and H&S. For those seven phenomena, the reliability attenuation
correction led to an increase in absolute ES heterogeneity $\tau$. For
the remaining five phenomena, estimates of ES heterogeneity were zero
before and after the correction procedure, leading to no change at all.
This leaves three phenomena where the attenuation correction did indeed
lead to a reduction of heterogeneity. However, it is crucial to note
that the hypothesis tests identified statistically significant
heterogeneity in only four out of 12 phenomena and for none of those a
reduction in heterogeneity took place. In all four cases of
statistically significant heterogeneity, the attenuation correction
procedure led to an increase in heterogeneity, contrary to the W&D and
H&S predictions.

To make sure that actual differences in score reliability exist,
heterogeneity in score reliability is assessed, making use of
Reliability Generalization Meta-Analysis (reference). @tbl-RGMA
summarises the results across all 12 phenomena. Overall, scales for
almost all phenomena appear have produced scores with an average score
reliability larger than .7. Only two phenomena (Albarracin_Priming_SAT
and Alter_Analytic_Processing) come with lower score reliability.
However, for those phenomena we could not identify statistically
significant heterogeneity in effect sizes in @tbl-meanES in the first
place. From the phenomena where the ES heterogeneity grew larger as a
result from the attenuation correction procedure, for six out of seven
phenomena, we observe statistically significant heterogeneity in score
reliability. Only the Shnabel_Willingness_Reconcile_Rev project produced
scores where no heterogeneity in score reliability was identified.
However, for this project, it is crucial to note that only a small
number of replications (8), was available for analysis. Since the power
of Cochran’s Q-test for heterogeneity largely depends on the number of
replications, this result is inconclusive.

@@

```{r, echo = FALSE}
#| label: tbl-RGMA
#| tbl-cap: Reliability Generalization Meta-Analysis
knitr::kable(x = read.csv(here("Tables/Results_RMA_alpha.csv")))
```

@@

$$\tau^2_{\delta_0} = [\mu_{\rho_{xx‘}}]^2 \tau^2_{\delta} + [\mu_{\delta}]^2 \tau^2_{\rho_{xx‘}}$$ {#eq-HS}

We adjusted @eq-HS using the notation used throughout this text. Here,
$\delta_0$ refers to the ES, undistorted by sampling error but not
corrected for measuring error, $\delta$ refers to the ES, undistorted by
sampling error and measurement error and ρxx‘ refers to score
reliability.

@eq-HS implies that heterogeneity in $\delta_0$ is essentially a
function where heterogeneity and mean value of $\delta$ are weighted by
mean score reliability ρxx‘ and its heterogeneity. While it may not be
self-evident in the equation itself, it is easy to construct cases where
heterogeneity in $\delta_0$ is inflated or deflated by heterogeneity in
ρxx‘, contradicting Hunter & Schmidt’s interpretation of this equation.
In @fig-hs, the results of employing @eq-HS to compute the heterogeneity
(variance) in corrected ES is highlighted. All else held equal, only the
individual level of mean reliability \$ \mu\*{\*\\rho{xx‘}}\$ is varied
from .5 to .8. Mean corrected ES $\mu_{\delta}$ is held constant at 1,
its variance $\tau^2_{\delta}$ held constant at .04 and the variance of
score reliability $\tau^2_{\rho_{xx‘}}$ held constant at .02.

```{r, echo = FALSE}
#| label: fig-HS
#| fig-cap: Implications by Hunter & Schmidt's @eg-HS

mu_delta <- 1
var_delta <- .2^2
mu_alpha <- seq(from = .5, to = .8, length.out = 100)
var_alpha <- .14^2

var_delta0 <- mu_alpha^2 * var_delta + mu_delta^2 * var_alpha
